---
title: "Project 3: demographic, Q_12, bright spots"
author: "Jian Quan Chen"
date: "2023-03-17"
output:
  html_document: default
  pdf_document: default
---

For this project, my group (Team 4: Gregg Maloy, Jacob Silver, Miguel Gomez, Umer Farooq) regularly met up on Google Meet to discuss how we were going to divide the work and share what we had coded so far. We found a dataset on Kaggle that surveyed about 24000 people with the objective of finding out "who is working with data, what’s happening with machine learning in different industries, and the best ways for new data scientists to break into the field." The dataset contains 43 questions presented in a wide format spanning 296 columns.

The first part of this markdown is going to be my contribution to the project, which was analyzing the demographic information, question 12 (programming language), and looking for the "bright spots" in data science. During one of our meetups, Jacob shared his code, which consisted of a function that took the columns we wanted as inputs and graphed them in a bar chart. His code also contained a function that graphed the cross between two questions. In my case, I used his function to determine which programming language various positions (data scientists, software engineers, etc.) used on a regular basis. At our final meetup, we shared what we had completed and sliced down which data would go in the presentation.

# Loading the libraries
```{r setup, include=FALSE}
library(tidyverse)
library(wordcloud)
library(forcats)
```

## Loading the data from Github
```{r}
df <- read.csv("https://raw.githubusercontent.com/LeJQC/MSDS/main/DATA%20607/Project%203/kaggle_survey_2022_responses.csv",na.strings = c("","N/A"))
```

## Checking out the data frame
```{r}
#There were 23998 respondents to the survey and 43 questions separated into 296 columns
dim(df)
#head(df[1:6],5)
```
## Age
```{r}
#Making a data frame with the Age of all the respondents 
df2 <- df[2]
age <- data.frame(df2[-1, ])
colnames(age) <- "Age"
```

### Ploting their age
The majority of the respondents were between age 18 to 29  
```{r}
age %>% 
  ggplot(aes(x=Age))+
  geom_bar()+
  coord_flip()+
  theme_minimal()
```
### Table of the age and percentage
```{r}
age_count <- age %>%  
  count(Age) %>% 
  mutate(percent = round(n/sum(n)*100)) %>% 
  arrange(desc(n))

knitr:: kable(age_count)
```

## Gender
```{r}
#Making a data frame with the Gender of all the respondents 
df3 <- df[3]
gender <- data.frame(df3[-1, ])
colnames(gender) <- "Gender"
```

### Count/percentage of the gender
Three fourths of the respondents were male
```{r}
gender_count <- gender %>% 
  count(Gender) %>% 
  mutate(percent = round(n/sum(n)*100)) %>% 
  arrange(desc(n)) 

knitr:: kable(gender_count)
```

## Country
```{r}
#Making a data frame with the Country of all the respondents 
df4 <- df[4]
country <- data.frame(df4[-1, ])
colnames(country) <- "Country"
```

### Count/percentage of the country
About a third of the respondents are from India
```{r}
country_per <- country %>% 
  count(Country) %>% 
  arrange(desc(n)) %>% 
  mutate(percent = round(n/sum(n)*100)) %>% 
  head(10) 

knitr:: kable(country_per)
```

# Question 12

## Programming Language
Looking at the most used common programming language (question 12) among the respondents
```{r}
#Splicing the columns needed
df12 <- df[,31:45]
```

### Converting the 15 columns of question 6 into long format
```{r}
dfq12 <- df12 %>% 
  pivot_longer(
    cols = everything(),
    names_to = "Question 12",
    values_to = "Language",
    values_drop_na = TRUE
  )

dfq12 <- dfq12[16:nrow(dfq12),]
```

### Plotting the info
```{r}
dfq12 %>% 
  ggplot(aes(x=Language))+
  geom_bar()+
  coord_flip()+
  theme_minimal()
```

### Tabling the data
```{r}
table(dfq12$Language)
```
## Question 12 with Question 11

### Creating a data frame to look at the relationship between years of experience writing code/programming(question 11) and the programming language use regularly(question 12)
```{r}
#Splicing the columns I want
df11 <- df[3:nrow(df),30:45]
```

### Cleaning out those who have no experience writing code
```{r}
df11_clean <- df11 %>% 
  filter(df11$Q11 != "I have never written code") %>%
  group_by(Q11) 
```

### Converting data to long format
```{r}
dfq11 <- df11_clean %>% 
  pivot_longer(
    cols = -1,
    names_to = "Question 12",
    values_to = "Language",
    values_drop_na = TRUE)
```

### Contingency table of years of experience and programming language
Seems python is the most used language regardless of years of experience
```{r}
table(dfq11$Q11,dfq11$Language)
```

### Plotting the data
It looks like the people just starting their data science career use more programming languages than data scientist who have more than 10 years of experience.
```{r}
dfq11 %>% 
  ggplot(aes(x= factor(Q11, levels = c("< 1 years","1-3 years","3-5 years","5-10 years","10-20 years","20+ years")),fill=Language))+
  geom_bar(alpha = 20)+
  theme_bw()+
  labs(title = "Years of Experience vs Language", x= "Years of Experience")
```

# Salary

## Creating a data frame of the compensation
```{r}
df_comp <- df[159]
comp <- na.omit(data.frame(df_comp[-1,]))
colnames(comp) <- c("Compensation")
```

### Just wanted to see what it looks like
```{r}
comp %>% 
  count(Compensation) %>% 
  arrange(n)
```

# "Bright Spots"-Top Earners 
Created a new data frame that filtered by people who made $200,000 or more. There were 380 respondents of the 23998 people survey who matched this criteria, which is about 1.6%. 
```{r}
df_top <- df %>% 
  filter(df$Q29 %in% c("200,000-249,999","250,000-299,999","300,000-499,999","$500,000-999,999",">$1,000,000"))

df_top <- df_top[,c(6:25,29:44,146)]
```

Instead of looking at each observation individually, I wanted to get a general sense of what these "bright spots" have in common. I dumped all the answers to questions 6(where they learned to code), 7(most useful platform), 12(programming language), and 23 (job title) into one data frame and made a wordcloud. 
```{r}
df_top_long <- df_top %>% 
  pivot_longer(
    cols = everything(),
    names_to = "Question",
    values_to = "Value",
    values_drop_na = TRUE)
```

```{r}
df_top_comp <- df_top_long %>%  
  count(Value) %>% 
  arrange(desc(n))

knitr:: kable(head(df_top_comp,10))
```

### Word Cloud of the answers
```{r}
df_top_long %>% 
  count(Value) %>% 
  with(wordcloud(Value, n, max.words = 10,colors = brewer.pal(10,"Spectral"),scale = c(2,.5)))
```

### This next portion is the code of the functions Jacob created. I used it to analyze the programming languaged use by different job titles and programming language in relation to years of experience and degree. 
```{r}
#clear question text from first row
df <- df[-1,]
```

## This function makes a bar chart of the column/question that is input
```{r}
graph_multicolumn_q <- function(og_df, q_list, 
                                q_name, val_label,
                                q_text) {
  #Create limited df with only that question's columns
  q_df <- df %>%
     select(all_of(q_list))
  
  #Create long-pivoted df for graphing
  q_df1 <- q_df %>%
    pivot_longer(
    cols = everything(),
    names_to = q_name,
    values_to = val_label,
    values_drop_na = TRUE)
  
  #Produce df of value counts
  q_df1_count <- q_df1 %>%
    count(!!sym(val_label))
  
  #Graph result
  
  q_df1 %>%
    ggplot(aes(x=fct_rev(fct_infreq(!!sym(val_label)))))+
    geom_bar()+
    coord_flip()+
    theme_minimal()+
    ggtitle(q_text)+
    xlab(val_label)
  
}
```

## This function graphs the questoin against another question
```{r}
graph_cross_analysis <- function(og_df, q_demo, q_list, q_name, val_label, q_text) {
  q_df <- og_df %>%
    select(all_of(c(q_demo, q_list)))
  
  q_df1 <- q_df %>%
    pivot_longer(
      cols= -1,
      names_to=q_name,
      values_to=val_label,
      values_drop_na=TRUE
  )
  
  q_df2 <- q_df1[complete.cases(q_df1),]  
    
  q_df2 %>%
    ggplot(aes(x=fct_rev(fct_infreq(!!sym(val_label))), fill = as.factor(!!sym(q_demo))))+
    facet_wrap(q_demo)+
    coord_flip()+
    geom_bar()+
    theme(axis.text.x = element_text(size = 7))+
    theme(axis.text.y = element_text(size = 7))+
    theme(legend.position = "none")+
    labs(y= "Count", x = val_label)+
    ggtitle(q_text)
    
    
}

```

## Creating a vector of the columns I want for question 12
```{r}
q12_cols <-  c("Q12_1","Q12_2","Q12_3","Q12_4","Q12_5", "Q12_6","Q12_7","Q12_8","Q12_9","Q12_10","Q12_11","Q12_12","Q12_13","Q12_14","Q12_15")
```

### Inputting this into the function to create a bar graph
```{r}
graph_multicolumn_q(df,
                    q12_cols,
                    'Question 12',
                    'Language',
                    'Regularly Used Programming Language')
```

### Crossing Language with Years of experience
```{r}
graph_cross_analysis(og_df = df, 
                     q_demo = 'Q11', 
                     q_list = q12_cols, 
                     q_name = 'Question 12', 
                     val_label = 'Language',
                     q_text = 'Language and Years of Experience')
```

### Crossing Language with Degree
```{r}
graph_cross_analysis(og_df = df, 
                     q_demo = 'Q8', 
                     q_list = q12_cols, 
                     q_name = 'Question 12', 
                     val_label = 'Language',
                     q_text = 'Language and Degree')
```

### Crossing Language with Job title
```{r}
graph_cross_analysis(og_df = df, 
                     q_demo = 'Q23', 
                     q_list = q12_cols, 
                     q_name = 'Question 12', 
                     val_label = 'Language',
                     q_text = 'Language and Job Title')
```