---
title: 'Week 7 Assignment: Working with XML and JSON in R'
author: "Jian Quan Chen"
date: "2023-03-12"
output:
  html_document: default
  pdf_document: default
---

For this week's assignment, I had to pick three books and store their information into a HTML, XML, and JSON format. Then, load the three files into R data frames. I can't remember the last time I picked up a book so I just picked three data science books I came across in the past year. Through several youtube videos, I created three files using Visual Studio Code and uploaded them to Github.

# Loading the Libraries

```{r setup, message=FALSE}
library(tidyverse)
library(rvest)
library(XML)
library(xml2)
library(jsonlite)
```

# Importing file from Github - HTML

```{r}
html <- read_html("https://raw.githubusercontent.com/LeJQC/MSDS/main/DATA%20607/Week%207%20Assignment/Week7%20books.html")
```

### Converting the HTML into a R dataframe

```{r}
#From the rvest library -> parsing a HTML table into a data frame
df_html <- html_table(html)
df_html <- df_html[[1]]
knitr:: kable(df_html)
```

# Importing file from Github - XML

```{r}
#From the xml2 library
xml_file <- read_xml("https://raw.githubusercontent.com/LeJQC/MSDS/main/DATA%20607/Week%207%20Assignment/Week%207%20books.xml")
```

### Converting XML to R data frame

```{r}
#Had to parse xml file before I could use the function xmltoDataFrame()
xml_parse <- xmlParse(xml_file)
#From the XML library
df_xml <- xmlToDataFrame(xml_parse)
knitr:: kable(df_xml)
```

# Importing file from Github - JSON

```{r}
json_file <- fromJSON("https://raw.githubusercontent.com/LeJQC/MSDS/main/DATA%20607/Week%207%20Assignment/week%207%20books.json")
```

### Converting JSON to R dataframe

```{r}
df_json <- as.data.frame(json_file)
knitr:: kable(df_json)
```

# Conclusion

All three data frames are pretty similar. They all have 3 rows with 5 columns and the data is exactly the same. The biggest difference that probably stands out is the column names, which I forgot to capitalize when I was creating the xml and json files. I changed the column names below and checked to see if they are identical.

```{r}
colnames(df_xml) <- colnames(df_html)
colnames(df_json) <- colnames(df_html)
identical(df_html,df_xml)
identical(df_html,df_json)
identical(df_xml,df_json)
```

It looks like the data frames from the xml and json files are identical. However, the data frame from the HTML seem to differ from the other two data frames. I used the all.equal() function to see if I could identify the difference and it looks like it is just a minor string mismatch.

```{r}
all.equal(df_html,df_xml)
all.equal(df_html,df_json)
```
