---
title: "Week 5 Assignment"
author: "Jian Quan Chen"
date: "2023-02-26"
output:
  html_document: default
  pdf_document: default
---
For this assignment, I had to transform a wide table into a long table and analyze the data. Converting the wide dataset of airline information to a long dataset was simple with the pivot_longer function in tidyr. Initially, I plotted the data to see if anything stood out. Afterwards, I mutated the dataframe to look at the percentage of delayed flights from each airline. Then I plotted the percentage of delayed flights by city to see if there were any trends.

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
```

#Created a csv in excel and uploaded it to github. Saved it as a dataframe(df)
```{r}
df <- read.csv("https://raw.githubusercontent.com/LeJQC/MSDS/main/DATA%20607/week%205%20flights.csv")

df
```

#Added some values and column names to df
```{r}
df[2,1] <- "ALASKA"
df[5,1] <- "AM WEST"
colnames(df) <- c("Airline", "Status", "Los Angeles", "Pheonix", "San Diego", "San Francisco", "Seattle")

df
```

#Used the pivot_longer function in tidyr to convert wide data into long data
```{r}
df_long <- df %>% 
  pivot_longer(
    cols = c("Los Angeles", "Pheonix", "San Diego", "San Francisco", "Seattle"),
    names_to = "City",
    values_to = "Value",
    values_drop_na = TRUE) %>% 
  select("Airline", "City", "Status", "Value") %>% 
  arrange(City, Status)

df_long
```

#Wanted to get a general sense of the data by plotting it. Looks like Phenoix has the largest amount of on time flights. 
```{r}
df_long %>% 
  ggplot(aes(x= City, y = Value, fill = Status))+
  facet_wrap(~Airline)+
  geom_bar(stat = "identity", position = "dodge")+
  labs(title = "Airline Delays", x = "City", y = "Count")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

###Analysis to compare the arrival delays for the two airlines

#Calculating the percentage of delayed flights on ALASKA airline
```{r}
df_delayed_alaska <- df_long %>% 
  filter(Airline == "ALASKA") %>% 
  mutate(total_flights = sum(Value)) %>% 
  filter(Status == "delayed") %>% 
  mutate(pct_flight_delay = sum(Value)/total_flights*100)

df_delayed_alaska
```

#Calculating the percentage of delayed flights on AM WEST airline
```{r}
df_delay_am <- df_long %>% 
  filter(Airline == "AM WEST") %>% 
  mutate(total_flights = sum(Value)) %>% 
  filter(Status == "delayed") %>% 
  mutate(pct_flight_delay = sum(Value)/total_flights*100)

df_delay_am
```

#Finding the percentage of delayed flights by city and airline
```{r}
df_delay_city <- df_long %>% 
  group_by(City,Airline) %>% 
  mutate(total_flights_city = sum(Value)) %>% 
  mutate(pct_delay_city = Value/total_flights_city*100)

df_delay_city
```

```{r}
df_delay_city %>% 
  filter(Status == "delayed") %>% 
  ggplot(aes(x= City, y = pct_delay_city, fill = Airline))+
  geom_bar(stat = "identity", position = "dodge")+
  labs(title = "Airline Delays by percentage", x = "City", y = "Percent Delayed")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Conclusion:
AM West airline had more flights at 7225 compared to Alaska Airlines at 3775. Even though they had almost double the number of flights as Alaska Airlines, AM West (10%) had a lower overall percentage of delayed flights than Alaska Airlines (13%).

However, if I were to compare the percentage of delays in each city, Alaska Airlines has a lower percentage of delayed flights in each city compared to AM West.

The reason AM West airline has a lower overall percentage of delayed flights is because Phoenix has almost 5,000 flights that were on time.